On the Functionality of the Multilayer Perceptron

### 2020-12-14 Thesis/Drafts/Buell-PHYS799-ThesisDraft3.tex
Minor updates to Thesis, New Citation

The task of recognizing the source of a sound wave in a digital audio file is trivial and almost effortless to a human being, but far more difficult for a computer to do. To execute this task, we have designed and constructed a multimodal neural network, that utilizes two different input branches, one being a convolutional neural network (CNN), and the other, a multilayer perceptron (MLP). The CNN branch accepts input in the form of a two-dimensional spectrogram matrix, which encodes the energy distribution of an audio signal over frequency and time, and the MLP branch accepts input as a one-dimensional array of features derived from the physical nature of the sound wave. The spectrogram is produced by using a concatenation of multiple short-time Fourier transformations (STFT) of overlapping windowed time-space analysis frames. The feature vector implements techniques from digital signal processing to combine more explicit properties of the time-space and frequency-space information. To classify the audio files based on predicted sources, we aggregate the of the activations from the final layers of each branch, and produce a single prediction label. We find that this architecture allows for reasonable classification performance, and demonstrates superior classification performance on average, when compared to either of it's constituent unimodal architectures. Training data for this experiment is from studio recordings of the Philharmonia Symphony Orchestra and University of Iowa's Electronic Music Studios.

@@

\paragraph*{}This project includes a great number of steps to ensure it's completion. Here we show a high-level outline of the pieces involved.

@@

\paragraph*{}To ensure that these data sets are formatted consistently, we read each sample from it's original format, \textit{.AIF}, \textit{.MP3}, or similar, and rewrite the data as a new \textit{.WAV} files, sampled at 44.1 kHz with a 24-bit depth. This ensures that all data will have a consistent format when features are extracted. This stage is detailed in sec. (\ref{subsubsec-Preprocessing}).

@@

\paragraph*{}The performance of a neural network is largely dependent of designing an appropriate set of predictors or features which represent the raw input given to the model. These inputs are designed to represent properties of wave forms that can be expressed by a single or multiple numerical values. We detail these features in sec. (\ref{sec-Features}). These are used in place of a full waveform to represent a file's contents in a low-dimensional array. We use tools from physics, mathematics, and signal processing to define and explore a comprehensive set of features that enables a high performance of the classifier.

@@

\paragraph*{}With the appropriate set of features designed, and the number of output categories determined, we can organize the structure of the neural network function. The shape of the network defines the flow of information, and the hypothesis space. The nature of all possible solutions that the trained model could arrive at is subject to the architecture \cite{Goodfellow,Loy}. We construct a multimodal neural network, explored in sec. (\ref{subsec-Architecture}), that processes two distinct entry points derived from the same audio sample. This network is designed to evaluate each input independently, and concatenate the results to produce a single outputted prediction.

@@

\paragraph*{}We execute a K-Folds cross validation algorithm to evaluate the performance of the neural network across overlapping subsets of data. This ensures that the model can perform consistently and reliably given small changes in initial conditions. In this stage, we also choose the value of hyper-parameters, activation functions and layer widths to best compliment the chosen features. This process is repeated and expanded upon until we have produced a model with a sufficient performance. We explore the explore this is sec. (\ref{sec-PerfEval}).

@@

\paragraph*{}Once we have established a suitable performance of the model, we allow the model to run predictions on the un-labeled chaotic synthesizer wave forms. We output the prediction results to a file, and compare the neural network predictions against human predictions. If further corrections are needed, we revert and re-deign the features, architecture, or hyper-parameters as needed. We explore thisas part of our conclusions in sec. (\ref{sec-Conclusion}).

@@

\paragraph*{}As the name implies, early neural networks were inspired from from the human brain. Former YouTube video classification team lead, and current machine learning consultant, Aurelien Geron writes about the relationship between biological brains and mathematical neural networks \cite{Geron}: 
\begin{quote}
Birds inspired us to fly, burdock plants inspired velcro, and nature has inspired many other inventions. It seems only logical then, to look to the brain's architecture for inspiration on how to build an intelligent machine.
\end{quote}
The result is a computer program that is reminiscent of the a brain in that it structured much like the brain, and can\textit{learn} from new data. Where a biological brain uses connections of chemical and electrical exchanges to process input and make decisions, the computational counterpart uses arithmetic and numerical exchanges to similarly process input and make decisions \cite{Levine,Bishop}. Where a brain uses physical connection of neurons through axons, the nueral network uses mathematical connections of neurons through mathematical operations \cite{Geron,Levine}.

@@

\paragraph*{}The neural network function $F^*$ uses a set of parameters $\Theta$, to map those inputs, $x$ to outputs, $y$ \cite{Goodfellow,James,Virtanen}. It is the process of updating the elements in the object $\Theta$ which allow us to train a neural network to perform the desired task. Similarly, it is the process of organizing and connecting neurons in the brain that allow humans to perform any specific task \cite{Geron,Levine}. Over the course of their lives, humans will be able to learn to match sounds to sources with relative ease, given the appropriate labels. In a similar manner, our neural network is presented a collection of predictors from sound files and a complementary set of labels indicating the source of each sound. This type of task is called \textit{classification}, which involves mapping the input data, $x$ to one of $k$ possible \textit{classes}, each represented by an integer, $0$ through $k-1$. Each of the $k$ classes represents a particular musical instrument or sound source that could have produced the sound-wave in the audio file \cite{Goodfellow,Loy,Virtanen}. In this section, we explore the motivation, structure, and training process of a neural network.

@@

\paragraph*{}We can denote this for $n$ arrays with function notation:
\begin{equation}
\label{eqn-ConcatenationFunction}
f^{(l)} : x^{(l-1)}_{a} \in \mathbb{R}^{(1 \times \alpha)} , x^{(l-1)}_{b} \in \mathbb{R}^{(1 \times \beta)} , ....  \rightarrow
x^{(l)}_{z} \in \mathbb{R}^{(1 \times \alpha + \beta + ...)}
\end{equation}
The concatenation layer is used to combine activations from two different layers into a single new layer. In this case of the model used in the project, we combine the outputs that result from the convolution branch and the perceptron branch to produce an single aggregated output. This process is detailed further in sec. (\ref{subsec-Architecture}). This layer also "non-trainable" because it does no use any weights or transformations, only a fixed procedure.

@@

\paragraph*{}For the categorical cross entropy cost function, we can evaluate $\partial J / \partial x^{(L-1)}$ over a single sample that belongs to class $i$ as:
\begin{equation}
\label{eqn-CXELossDeriv}
\frac{\partial J}{\partial x^{(L-1)}} = \frac{\partial J}{\partial y^*} =
\frac{\partial}{\partial y^*}\Big[ -y \ln (y^*) \Big] = -\frac{y}{y^*_i + \delta} = -\frac{1}{y^*_i + \delta}
\end{equation}
We often introduce $\delta \approx 10^{-8}$ to prevent possible division by $0$ errors. The value $\partial x^{(L-1)} / \partial z^{(L-1)}$ can be found by differentiating Eq. (\ref{eqn-elementActivation}) with respect to the input activations.
\begin{equation}
\frac{\partial x^{(L-1)}}{\partial z^{(L-1)}} = \sigma^{'(l)} \big[ z^{(L-1)} \big]
\end{equation}
For a dense layer as in sec. (\ref{subsubsec-DenseLayer}) or a 2D-convolution layer as in sec. (\ref{subsubsec-Conv2DLayer}) $\partial z^{(L-1)} / \partial W^{(L-1)}$ is given by the derivative of Eq. (\ref{eqn-LinearTransform}) or Eq. (\ref{eqn-LinearTransform2}) with respect to $W^{(l)}$:
\begin{equation}
\frac{\partial z^{(L-1)}}{\partial W^{(L-1)}} = 
\frac{\partial}{W^{(L-1)}}\Big[ W^{(L-1)}x^{(L-2)} + b^{(L-1)}\Big] = x^{(L-2)}
\end{equation}
\begin{equation}
\frac{\partial z^{(L-1)}}{\partial W^{(L-1)}} = 
\frac{\partial}{W^{(L-1)}}\Big[ \big( W^{(L-1)} * x^{(L-2)} \big) + b^{(L-1)}\Big] = x^{(L-2)}
\end{equation}
Finally, differentiating with respect to the bias array: $\partial z^{(L-1)} / \partial b^{(L-1)}$, we have:
\begin{equation}
\frac{\partial z^{(L-1)}}{\partial b^{(L-1)}} = \frac{\partial}{W^{(L-1)}}\Big[ W^{(L-1)}x^{(L-2)} + b^{(L-1)}\Big] = 1
\end{equation}

@@

\paragraph*{}An optimizer is the algorithm or procedure that is used by a machine learning model to perform the optimization task to reduce the cost value after reciveing each new training sample. Regression models may often used a \textit{mean-squared error} cost function which can often be solved analytically, while high dimensional neural networks, such as the one used in this project, outlined in sec. (\ref{subsec-Architecture}), can only up updated iterativly \cite{Goodfellow,James,Loy}. We use a variation of gradient descent much like the one in Eq. (\ref{eqn-GradientLearning}). A standard stochastic gradient-based or batch gradient descent learning algorithm can be fast, but are often prone errors such as vanishing or exploding gradients \cite{Geron,Goodfellow,Loy}. To combat this, model must often implement an optimizer that is more stable, and robust in it's ability to drive the cost function to a successively lower value.

@@

\paragraph*{}As the optimizer iterates through each step, the effect of past iterations will tend to "snowball". If the previous step was found to reduce the cost function by a large or small amount, the step size for the previous step will update accordingly. This enables to ADAM to overcome smaller discontunities that may arise in the solution space \cite{Goodfellow}. For a given step $(i)$, The ADAM update is given:
\begin{equation}
\label{eqn-ADAMupdate}
\begin{split}
s^{(i)} &= \rho_1 s^{(i-1)} + (1 - \rho_1) \nabla_{\Theta} \bar{J}  \\
r^{(i)} &= \rho_2 r^{(i-1)} + (1 - \rho_2) \Big[ \nabla_{\Theta} \bar{J} \odot \nabla_{\Theta} \bar{J} \Big]  \\
s'^{(i)} &= \frac{s^{(i)}}{1-\rho_1^i} \\
r'^{(i)} &= \frac{r^{(i)}}{1-\rho_2^i} \\
\Theta^{(i)} &= \Theta^{(i-1)} + (-\alpha)\frac{s'^{(i)}}{\sqrt{r'^{(i)}}+\delta} \\
\end{split}
\end{equation}
Note that a superscript $(i)$ or $(i-1)$ gives an iteration index, while the superscript $i$ means to raise a value to the power of $i$. ADAM has experimentally shown to be a very powerful and robust optimizer useful for a wide range of tasks \cite{Goodfellow}. Because of the two decay constants $\rho_1$ and $\rho_2$, we can compound and accumulate the values of past gradients to continue to push the cost to lower and lower values, even if the magnitude of the gradient becomes very small \cite{Geron}.

@@

\paragraph*{}We can carry this idea over by producing a \textit{multimodal} neural network that accepts two different forms of input for each sample. One input is an $N' \times k$ spectrogram matrix, explored in sec. (\ref{subsec-Spectrogram}) and the other input is a $1 \times p$ feature vector explored in sec. (\ref{subsec-TimeFeatures}) and sec. (\ref{subsec-FreqFeatures}). The spectrogram acts to provide an energy distribution of a signal as a function of frequency and time, while the feature vector provides qualities in a more list-like fashion. This type of supervised learning is referred to as \textit{multi-representation learning}, and we detail a conceptual diagram of the architecture in Fig. (\ref{fig-NetworkArchitecture}) \cite{Li,Ngiam}.

@@

\paragraph*{}The input layer holds the input activations, $X_1$, which is a collection of $64$ spectrograms. The output of the the input layer is passed into the first of three \textit{Convolution Layer Groups}. These layer groups are inspired from the \textit{Visual Geometry Group-16 Neural Network} architecture \cite{Goodfellow,Loy}. Each convolution layer group is composed of three individual layers: 
\begin{enumerate}
\item A 2-Dimensional Convolution layer, $32$ filters, $3 \times 3$ kernel, $1 \times 1$ step size, ReLU activation function,
\item A 2-Dimensional Convolution layer, $32$ filters, $3 \times 3$ kernel, $1 \times 1$ step size, ReLU activation function,
\item A 2-Dimensional Maximum Pooling layer, $3 \times 3$ pooling size, Identity activation function
\end{enumerate}  
The convolution layers iterate over the middle two axes of the data, see sec. (\ref{subsubsec-Conv2DLayer}). Physically, this allows us to search through the frequency and time axes to detect dominant features or shapes.

@@

\paragraph*{}The perceptron branch is pictured on the right side of Fig. (\ref{fig-NetworkArchitecture}) and is an implementation of a \textit{multilayer percepton neural network} (MLP). Rather than accept an image-like input, the perception simply takes a $p$-dimensional vector-like input of properties that are derived from the audio file. We call these properties \textit{features} or \textit{predictors} \cite{Geron,Khan,Serizel}. Details on the natures and selections of these $p$ elements are found in sec. (\ref{subsec-TimeFeatures}) and sec. (\ref{subsec-FreqFeatures}).

@@

\paragraph*{}In perceptron models, the scaling of the design matrix is vital to classification performance. A design matrix is scaled by taking all samples in a column (a particular feature from each class), subtracting the average, and then scaling it such that it has unit variance \cite{Geron,James}. This ensures that no one feature dominates the performance of the model and that layer activations do not get saturated to extremes as data progresses through the network.

@@

\paragraph*{}The last layer in each of the two branches is a ReLU-activated Dense layer containing 64 neurons, represented by a $64$ - dimensional vector. We combine these model layers by using a \textit{concatenation layer} to fuse the arrays together, such that each sample is given by a $128$-dimensional vector. This vector is then transformed into the final dense layer, which uses the softmax activation, and encodes the joint predictions based on contributions of both model branches. In sec. (\ref{sec-Results}) we explore how the predictive power of the network changes when examining the performance of each individual branch compared to the hybrid model.

@@

\paragraph*{}As discussed in Sec. (\ref{sec-TheNeuralNetwork}), performance quality of a neural network is greatly dependent on the properties of the chosen features \cite{Virtanen,Liu}. To ensure that the classifier model performs adequately and consistently, we must choose these features to have high variance between classes, and low variance within each class. This enables a neural network to development clear decision boundaries between each unique class \cite{James,Serizel}. To develop this set of features, we explore the physical and mechanical properties of musical instruments and the sounds that they create.

\paragraph*{}Since computer algorithms such are neural networks are built to handle exclusively quantitative information, we must ensure that each property can be represented by a numerical quantity \cite{Geron,Levine}. Where a human would use qualitative adjectives such as \textit{bright} or \textit{harsh} or \textit{percussive} to describe sound, computers require a numerical descriptor instead. The goal of this section is to develop the attributes of sound that enable us to describe a sound wave in a quantitative and compact way. Thus we use values that can be expressed as numbers such as \textit{amplitude envelopes}, \textit{frequency spectra}, or \textit{formant structure}. For this section, we represent waveforms as \textit{spectrograms} to capture the time-space and frequency-space development of energy simultaneously. We motivate and describe the production of a spectrogram in sec. (\ref{subsec-Spectrogram}).

@@

\paragraph*{}We model the tight elastic membrane very similarly as that of a tight elastic string, except in two dimensions \cite{White}. Unlike the chordophone sec. (\ref{subsec-chordophones}), the partials for the membrane are not related by integer multiples. This lack of harmonic structure leaves most membranophones to have a sort of "muddled" frequency spectrum where we see a more continuous distribution of constituent frequencies arise as opposed to clean integer related overtones \cite{White,Olson}. This partly why many membranophones have a much more "percussive" sound as opposed to performing any distinct note.

@@

\paragraph*{}From each audio file sample, we calculate the values for the same predictors. The predictors are arranged into two separated structures, each called a \textit{design matrix}, described in sec. (\ref{subsubsec-DesignMatrix}) \cite{Goodfellow,Geron}. For a model with $p$ predictors, feature space is a $p$-dimensional vector space where each basis vector represents a predictor \cite{Goodfellow,James}. Each sample can be described by a $p$-dimensional vector, given by each row of the design matrix $X_2$. For samples $x^{(a)}$ and $x^{(b)}$ that belong to the same class, we expect the vector to be similar, $x^{(a)} \approx x^{(b)}$ in some way. This indicates that the two samples lie "close together" in space. Similarly, for samples $x^{(c)}$ and $x^{(d)}$ that belong to different classes, we expect the vectors to be very different, indicating that the two sample lie "far apart" in space. Note that this is an idealization, and not always the defining case for the decision function behavior.

@@

\paragraph*{}Each audio sample is represented by an $N' \times k$ spectrogram matrix explored in sec. (\ref{subsec-Spectrogram}), a $1 \times p$ feature vector, and a  one-hot-encoded target label, $y$. In order to efficiently present the information to the model for training or predicting, we extract a batch of $b$ samples and produce the two arrays form each. Since the spectrogram and feature vector represent different \textit{modalities}, we construct two separate \textit{design matrices}, $X_1$ and $X_2$ \cite{Geron,Li,Ngiam}. Each has shape given by:
\begin{equation}
\label{eqn-X1Shape}
X_1 \in \mathbb{R}^{(b \times N' \times k)}
\end{equation}
\begin{equation}
\label{eqn-X2Shape}
X_2 \in \mathbb{R}^{(b \times p)}
\end{equation}
Where $N' \times k$ gives the shape of the spectrogram matrix $S$ from a single sample, described in sec. (\ref{subsec-Spectrogram}). Similarly, $p$ is the number of predictors extracted from each sample, outlined in sec. (\ref{subsec-TimeFeatures}) and sec.(\ref{subsec-FreqFeatures}). The first axis, with size $b$ indicates that there are $b$ audio file samples stored in that design matrix. Organizing the samples in this fashion allows for easy machine and human interpretation of each axis and index.

@@

\paragraph*{}In addition to the standard confusion matrix, we can also weight entries by prediction score. Rather than adding $+1$ to each entry $C_{i,j}$ where appropriate, we instead add the \textit{probability} value, which is bounded $[0,1]$ (See softmax activation function in sec. \ref{subsubsec-Softmax})). This allows us to bake in the predictive \textit{confidence} into the confusion matrix. For example, two correct predictions with a score $0.2$ are weighted equally to a single incorrect prediction of $0.4$. This is useful for identifying a classifier that makes correct predictions, but only by smaller threshold. This can almost be through of as assigning "partial credit" to each prediction.

>>>>

The task of recognizing the source of a sound wave in a digital audio file is trivial and almost effortless to a human being, but far more difficult for a computer to do. To execute this task, we have designed and constructed a multimodal neural network, that utilizes two different input branches, one being a convolutional neural network (CNN), and the other, a multilayer perceptron (MLP). The CNN branch accepts input in the form of a two-dimensional spectrogram matrix, which encodes the energy distribution of an audio signal over frequency and time, and the MLP branch accepts input as a one-dimensional array of features derived from the physical nature of the sound wave. The spectrogram is produced by using a concatenation of multiple short-time Fourier transformations (STFT) of overlapping windowed time-space analysis frames. The feature vector implements techniques from digital signal processing to combine more explicit properties of the time-space and frequency-space information. To classify the audio files based on predicted sources, we aggregate the of the activations from the final layers of each branch, and produce a single prediction label. We find that this architecture allows for reasonable classification performance, and demonstrates superior classification performance on average, when compared to either of it's constituent unimodal architectures. Training data for this experiment is from studio recordings of the Philharmonia Symphony Orchestra and the University of Iowa's Electronic Music Studios.

@@

\paragraph*{}This project includes a number of steps to ensure it's completion. Here we show a high-level outline of the pieces involved.

@@

\paragraph*{}To ensure that these data sets are formatted consistently, we read each sample from it's original format, \textit{.AIF}, \textit{.MP3}, or similar, and rewrite the data as a new \textit{.WAV} files, sampled at 44.1 kHz with a 24-bit depth. This ensures that all data will have a consistent format when features are extracted. This stage is detailed in section (\ref{subsubsec-Preprocessing}).

@@

\paragraph*{}The performance of a neural network is largely dependent of designing an appropriate set of predictors or features which represent the raw input given to the model. These inputs are designed to represent properties of wave forms that can be expressed by a single or multiple numerical values. We detail these features in section (\ref{sec-Features}). These are used in place of a full waveform to represent a file's contents in a low-dimensional array. We use tools from physics, mathematics, and signal processing to define and explore a comprehensive set of features that enables a high performance of the classifier.

@@

\paragraph*{}With the appropriate set of features designed, and the number of output categories determined, we can organize the structure of the neural network function. The shape of the network defines the flow of information, and the hypothesis space. The nature of all possible solutions that the trained model could arrive at is subject to the architecture \cite{Goodfellow,Loy}. We construct a multimodal neural network, explored in section (\ref{subsec-Architecture}), that processes two distinct entry points derived from the same audio sample. This network is designed to evaluate each input independently, and concatenate the results to produce a single outputted prediction.

@@

\paragraph*{}We execute a K-Folds cross validation algorithm to evaluate the performance of the neural network across overlapping subsets of data. This ensures that the model can perform consistently and reliably given small changes in initial conditions. In this stage, we also choose the value of hyper-parameters, activation functions and layer widths to best compliment the chosen features. This process is repeated and expanded upon until we have produced a model with a sufficient performance. We explore the explore this is section (\ref{sec-PerfEval}).

@@

\paragraph*{}Once we have established a suitable performance of the model, we allow the model to run predictions on the un-labeled chaotic synthesizer wave forms. We output the prediction results to a file, and compare the neural network predictions against human predictions. If further corrections are needed, we revert and re-deign the features, architecture, or hyper-parameters as needed. We explore thisas part of our conclusions in section (\ref{sec-Conclusion}).

@@

\paragraph*{}As the name implies, early neural networks were inspired from from the human brain. Former YouTube video classification team lead, and current machine learning consultant, Aurelien Geron writes about the relationship between biological brains and mathematical neural networks \cite{Geron}: 
\begin{quote}
Birds inspired us to fly, burdock plants inspired velcro, and nature has inspired many other inventions. It seems only logical, then, to look to the brain's architecture for inspiration on how to build an intelligent machine.
\end{quote}
The result is a computer program that is reminiscent of the a brain in that it structured much like the brain, and can\textit{learn} from new data. Where a biological brain uses connections of chemical and electrical exchanges to process input and make decisions, the computational counterpart uses arithmetic and numerical exchanges to similarly process input and make decisions \cite{Levine,Bishop}. Where a brain uses physical connection of neurons through axons, the neural network uses mathematical connections of neurons through mathematical operations \cite{Geron,Levine}.

@@

\paragraph*{}The neural network function $F^*$ uses a set of parameters $\Theta$, to map those inputs, $x$ to outputs, $y$ \cite{Goodfellow,James,Virtanen}. It is the process of updating the elements in the object $\Theta$ which allow us to train a neural network to perform the desired task. Similarly, it is the process of organizing and connecting neurons in the brain that allow humans to perform any specific task \cite{Geron,Levine}. Over the course of their lives, humans will be able to learn to match sounds to sources with relative ease, given the appropriate labels. In a similar manner, our neural network is presented a collection of predictors from sound files and a complementary set of labels indicating the source of each sound. This type of task is called \textit{classification}, which involves mapping the input data, $x$ to one of $k$ possible \textit{classes}, each represented by an integer, $0$ through $k-1$. Each of the $k$ classes represents a particular musical instrument or sound source that could have produced the sound-wave in the audio file \cite{Goodfellow,Loy,Virtanen}. In this section, we explore the motivation, structure, and training process of a neural network. We implement our neural network with the \textit{Tensorflow} Python package \cite{Tensorflow}.

@@

\paragraph*{}We can denote this for $n$ arrays with function notation:
\begin{equation}
\label{eqn-ConcatenationFunction}
f^{(l)} : x^{(l-1)}_{a} \in \mathbb{R}^{(1 \times \alpha)} , x^{(l-1)}_{b} \in \mathbb{R}^{(1 \times \beta)} , ....  \rightarrow
x^{(l)}_{z} \in \mathbb{R}^{(1 \times \alpha + \beta + ...)}
\end{equation}
The concatenation layer is used to combine activations from two different layers into a single new layer. In this case of the model used in the project, we combine the outputs that result from the convolution branch and the perceptron branch to produce an single aggregated output. This process is detailed further in section (\ref{subsec-Architecture}). This layer also "non-trainable" because it does no use any weights or transformations, only a fixed procedure.

@@

\paragraph*{}For the categorical cross entropy cost function, we can evaluate $\partial J / \partial x^{(L-1)}$ over a single sample that belongs to class $i$ as:
\begin{equation}
\label{eqn-CXELossDeriv}
\frac{\partial J}{\partial x^{(L-1)}} = \frac{\partial J}{\partial y^*} =
\frac{\partial}{\partial y^*}\Big[ -y \ln (y^*) \Big] = -\frac{y}{y^*_i + \delta} = -\frac{1}{y^*_i + \delta}
\end{equation}
We often introduce $\delta \approx 10^{-8}$ to prevent possible division by $0$ errors. The value $\partial x^{(L-1)} / \partial z^{(L-1)}$ can be found by differentiating Eq. (\ref{eqn-elementActivation}) with respect to the input activations.
\begin{equation}
\frac{\partial x^{(L-1)}}{\partial z^{(L-1)}} = \sigma^{'(l)} \big[ z^{(L-1)} \big]
\end{equation}
For a dense layer as in section (\ref{subsubsec-DenseLayer}) or a 2D-convolution layer as in section (\ref{subsubsec-Conv2DLayer}) $\partial z^{(L-1)} / \partial W^{(L-1)}$ is given by the derivative of Eq. (\ref{eqn-LinearTransform}) or Eq. (\ref{eqn-LinearTransform2}) with respect to $W^{(l)}$:
\begin{equation}
\frac{\partial z^{(L-1)}}{\partial W^{(L-1)}} = 
\frac{\partial}{W^{(L-1)}}\Big[ W^{(L-1)}x^{(L-2)} + b^{(L-1)}\Big] = x^{(L-2)}
\end{equation}
\begin{equation}
\frac{\partial z^{(L-1)}}{\partial W^{(L-1)}} = 
\frac{\partial}{W^{(L-1)}}\Big[ \big( W^{(L-1)} * x^{(L-2)} \big) + b^{(L-1)}\Big] = x^{(L-2)}
\end{equation}
Finally, differentiating with respect to the bias array: $\partial z^{(L-1)} / \partial b^{(L-1)}$, we have:
\begin{equation}
\frac{\partial z^{(L-1)}}{\partial b^{(L-1)}} = \frac{\partial}{W^{(L-1)}}\Big[ W^{(L-1)}x^{(L-2)} + b^{(L-1)}\Big] = 1
\end{equation}

@@

\paragraph*{}An optimizer is the algorithm or procedure that is used by a machine learning model to perform the optimization task to reduce the cost value after reciveing each new training sample. Regression models may often used a \textit{mean-squared error} cost function which can often be solved analytically, while high dimensional neural networks, such as the one used in this project, outlined in section (\ref{subsec-Architecture}), can only up updated iteratively \cite{Goodfellow,James,Loy}. We use a variation of gradient descent much like the one in Eq. (\ref{eqn-GradientLearning}). A standard stochastic gradient-based or batch gradient descent learning algorithm can be fast, but are often prone errors such as vanishing or exploding gradients \cite{Geron,Goodfellow,Loy}. To combat this, model must often implement an optimizer that is more stable, and robust in it's ability to drive the cost function to a successively lower value.

@@

\paragraph*{}As the optimizer iterates through each step, the effect of past iterations will tend to "snowball". If the previous step was found to reduce the cost function by a large or small amount, the step size for the previous step will update accordingly. This enables to ADAM to overcome smaller discontinuities that may arise in the model's solution space \cite{Goodfellow}. For a given step $(i)$, The ADAM update is given:
\begin{equation}
\label{eqn-ADAMupdate}
\begin{split}
s^{(i)} &= \rho_1 s^{(i-1)} + (1 - \rho_1) \nabla_{\Theta} \bar{J}  \\
r^{(i)} &= \rho_2 r^{(i-1)} + (1 - \rho_2) \Big[ \nabla_{\Theta} \bar{J} \odot \nabla_{\Theta} \bar{J} \Big]  \\
s'^{(i)} &= \frac{s^{(i)}}{1-\rho_1^i} \\
r'^{(i)} &= \frac{r^{(i)}}{1-\rho_2^i} \\
\Theta^{(i)} &= \Theta^{(i-1)} + (-\alpha)\frac{s'^{(i)}}{\sqrt{r'^{(i)}}+\delta} \\
\end{split}
\end{equation}
Note that a superscript $(i)$ or $(i-1)$ gives an iteration index, while the superscript $i$ means to raise a value to the power of $i$. ADAM has experimentally shown to be a very powerful and robust optimizer useful for a wide range of tasks \cite{Goodfellow}. Because of the two decay constants $\rho_1$ and $\rho_2$, we can compound and accumulate the values of past gradients to continue to push the cost to lower and lower values, even if the magnitude of the gradient becomes very small \cite{Geron}.

@@

\paragraph*{}We can carry this idea over by producing a \textit{multimodal} neural network that accepts two different forms of input for each sample. One input is an $N' \times k$ spectrogram matrix, explored in section (\ref{subsec-Spectrogram}) and the other input is a $1 \times p$ feature vector explored in section (\ref{subsec-TimeFeatures}) and section (\ref{subsec-FreqFeatures}). The spectrogram acts to provide an energy distribution of a signal as a function of frequency and time, while the feature vector provides qualities in a more list-like fashion. This type of supervised learning is referred to as \textit{multi-representation learning}, and we detail a conceptual diagram of the architecture in Fig. (\ref{fig-NetworkArchitecture}) \cite{Li,Ngiam}.

@@

\paragraph*{}The input layer holds the input activations, $X_1$, which is a collection of $64$ spectrograms. The output of the the input layer is passed into the first of three \textit{Convolution Layer Groups}. These layer groups are inspired from the \textit{Visual Geometry Group-16 Neural Network} architecture \cite{Goodfellow,Loy}. Each convolution layer group is composed of three individual layers: 
\begin{enumerate}
\item A 2-Dimensional Convolution layer, $32$ filters, $3 \times 3$ kernel, $1 \times 1$ step size, ReLU activation function,
\item A 2-Dimensional Convolution layer, $32$ filters, $3 \times 3$ kernel, $1 \times 1$ step size, ReLU activation function,
\item A 2-Dimensional Maximum Pooling layer, $3 \times 3$ pooling size, Identity activation function
\end{enumerate}  
The convolution layers iterate over the middle two axes of the data, see section (\ref{subsubsec-Conv2DLayer}). Physically, this allows us to search through the frequency and time axes to detect dominant features or shapes.

@@

\paragraph*{}The perceptron branch is pictured on the right side of Fig. (\ref{fig-NetworkArchitecture}) and is an implementation of a \textit{multilayer percepton neural network} (MLP). Rather than accept an image-like input, the perception simply takes a $p$-dimensional vector-like input of properties that are derived from the audio file. We call these properties \textit{features} or \textit{predictors} \cite{Geron,Khan,Serizel}. Details on the natures and selections of these $p$ elements are found in section (\ref{subsec-TimeFeatures}) and section (\ref{subsec-FreqFeatures}).

@@

\paragraph*{}In perceptron models, the scaling of the design matrix is vital to classification performance. A design matrix is scaled by taking all samples in a column (a particular feature from each class), subtracting the average, and then scaling it such that it has unit variance \cite{Geron,James}. This ensures that no one feature dominates the performance of the model and that layer activations do not get saturated to extremes as data progresses through the network. This operation was implemented through the \textit{scikit-learn} Python package \cite{Sklearn}.

@@

\paragraph*{}The last layer in each of the two branches is a ReLU-activated Dense layer containing 64 neurons, represented by a $64$ - dimensional vector. We combine these model layers by using a \textit{concatenation layer} to fuse the arrays together, such that each sample is given by a $128$-dimensional vector. This vector is then transformed into the final dense layer, which uses the softmax activation, and encodes the joint predictions based on contributions of both model branches. In section (\ref{sec-Results}) we explore how the predictive power of the network changes when examining the performance of each individual branch compared to the hybrid model.

@@

\paragraph*{}As discussed in section (\ref{sec-TheNeuralNetwork}), performance quality of a neural network is greatly dependent on the properties of the chosen features \cite{Virtanen,Liu}. To ensure that the classifier model performs adequately and consistently, we must choose these features to have high variance between classes, and low variance within each class. This enables a neural network to development clear decision boundaries between each unique class \cite{James,Serizel}. To develop this set of features, we explore the physical and mechanical properties of musical instruments and the sounds that they create.

\paragraph*{}Since computer algorithms such are neural networks are built to handle exclusively quantitative information, we must ensure that each property can be represented by a numerical quantity \cite{Geron,Levine}. Where a human would use qualitative adjectives such as \textit{bright} or \textit{harsh} or \textit{percussive} to describe sound, computers require a numerical descriptor instead. The goal of this section is to develop the attributes of sound that enable us to describe a sound wave in a quantitative and compact way. Thus we use values that can be expressed as numbers such as \textit{amplitude envelopes}, \textit{frequency spectra}, or \textit{formant structure}. For this section, we represent waveforms as \textit{spectrograms} to capture the time-space and frequency-space development of energy simultaneously. We motivate and describe the production of a spectrogram in section (\ref{subsec-Spectrogram}).

@@

\paragraph*{}We model the tight elastic membrane very similarly as that of a tight elastic string, except in two dimensions \cite{White}. Unlike the chordophone section (\ref{subsec-chordophones}), the partials for the membrane are not related by integer multiples. This lack of harmonic structure leaves most membranophones to have a sort of "muddled" frequency spectrum where we see a more continuous distribution of constituent frequencies arise as opposed to clean integer related overtones \cite{White,Olson}. This partly why many membranophones have a much more "percussive" sound as opposed to performing any distinct note.

@@

\paragraph*{}From each audio file sample, we calculate the values for the same predictors. The predictors are arranged into two separated structures, each called a \textit{design matrix}, described in section (\ref{subsubsec-DesignMatrix}) \cite{Goodfellow,Geron}. For a model with $p$ predictors, feature space is a $p$-dimensional vector space where each basis vector represents a predictor \cite{Goodfellow,James}. Each sample can be described by a $p$-dimensional vector, given by each row of the design matrix $X_2$. For samples $x^{(a)}$ and $x^{(b)}$ that belong to the same class, we expect the vector to be similar, $x^{(a)} \approx x^{(b)}$ in some way. This indicates that the two samples lie "close together" in space. Similarly, for samples $x^{(c)}$ and $x^{(d)}$ that belong to different classes, we expect the vectors to be very different, indicating that the two sample lie "far apart" in space. Note that this is an idealization, and not always the defining case for the decision function behavior.

@@

\paragraph*{}Each audio sample is represented by an $N' \times k$ spectrogram matrix explored in section (\ref{subsec-Spectrogram}), a $1 \times p$ feature vector, and a  one-hot-encoded target label, $y$. In order to efficiently present the information to the model for training or predicting, we extract a batch of $b$ samples and produce the two arrays form each. Since the spectrogram and feature vector represent different \textit{modalities}, we construct two separate \textit{design matrices}, $X_1$ and $X_2$ \cite{Geron,Li,Ngiam}. Each has shape given by:
\begin{equation}
\label{eqn-X1Shape}
X_1 \in \mathbb{R}^{(b \times N' \times k)}
\end{equation}
\begin{equation}
\label{eqn-X2Shape}
X_2 \in \mathbb{R}^{(b \times p)}
\end{equation}
Where $N' \times k$ gives the shape of the spectrogram matrix $S$ from a single sample, described in section (\ref{subsec-Spectrogram}). Similarly, $p$ is the number of predictors extracted from each sample, outlined in section (\ref{subsec-TimeFeatures}) and section(\ref{subsec-FreqFeatures}). The first axis, with size $b$ indicates that there are $b$ audio file samples stored in that design matrix. Organizing the samples in this fashion allows for easy machine and human interpretation of each axis and index.

@@

\paragraph*{}In addition to the standard confusion matrix, we can also weight entries by prediction score. Rather than adding $+1$ to each entry $C_{i,j}$ where appropriate, we instead add the \textit{probability} value, which is bounded $[0,1]$ (See softmax activation function in section \ref{subsubsec-Softmax})). This allows us to bake in the predictive \textit{confidence} into the confusion matrix. For example, two correct predictions with a score $0.2$ are weighted equally to a single incorrect prediction of $0.4$. This is useful for identifying a classifier that makes correct predictions, but only by smaller threshold. This can almost be through of as assigning "partial credit" to each prediction.
